The Werewolf Among Us: Humans vs LLMs in Multi-Agent Games

### 2025-05-07 _tex/paper.tex
Built site for gh-pages

\subsection{Multi-Agent LLMs}\label{multi-agent-llms}

\begin{itemize}
\tightlist
\item
  Among us game (\citeproc{ref-chiAMONGAGENTSEvaluatingLarge2024}{Chi,
  Mao, and Tang 2024})
\item
  Collective problem solving
  (\citeproc{ref-duLargeLanguageModels2024}{Du, Rajivan, and Gonzalez
  2024})

  \begin{itemize}
  \tightlist
  \item
    ``analyses indicate that LLM agent groups exhibit more
    disagreements, complex statements, and a propensity for positive
    statements compared to human groups''
  \end{itemize}
\item
  Govsim (\citeproc{ref-piattiCooperateCollapseEmergence2024}{Piatti et
  al. 2024})

  \begin{itemize}
  \tightlist
  \item
    ``In GOVSIM, a society of AI agents must collectively balance
    exploiting a common resource with sustaining it for future use. This
    environment enables the study of how ethical considerations,
    strategic planning, and negotiation skills impact cooperative
    outcomes.''
  \end{itemize}
\item
  All found similar themes

  \begin{itemize}
  \tightlist
  \item
    That LLMs are capable and good at understanding the rules
  \item
    That they can cooperate and be sneaky
  \end{itemize}
\end{itemize}

\subsection{LLMs and Werewolf}\label{llms-and-werewolf}

\begin{itemize}
\tightlist
\item
  Examination of improving werewolf by LLMs
  (\citeproc{ref-xuLanguageAgentsReinforcement2024}{Xu et al. 2024})

  \begin{itemize}
  \tightlist
  \item
    ``our agents use an LLM to perform deductive reasoning and generate
    a diverse set of action candidates. Then an RL policy trained to
    optimize the decision-making ability chooses an action from the
    candidates to play in the game. Extensive experiments show that our
    agents overcome the intrinsic bias and outperform existing LLM-based
    agents in the Werewolf game.''
  \end{itemize}
\item
  Werewolf Arena (\citeproc{ref-bailisWerewolfArenaCase2024}{Bailis,
  Friedhoff, and Chen 2024})

  \begin{itemize}
  \tightlist
  \item
    Used in this paper
  \end{itemize}
\item
  Explicitly discuss how none of the exisiting LLM+Werewolf papers
  examine the differences/compare from a human dataset
\end{itemize}

@@

Agents participated in gameplay through a graphical user interface
(GUI), depicted in Figure Figure~\ref{fig-wwa-gui}. The GUI displays the
game state, including player roles, actions, and current discussion
rounds, enabling monitoring of the gameplay progression.

\textbf{Figure 1.} GUI of Werewolf Arena simulation environment.

>>>>

Recent research into multi-agent large language models (LLMs) has
explored their performance in various social deduction and collective
problem-solving contexts. Chi et al.~{[}Chi, Mao, and Tang
(\citeproc{ref-chiAMONGAGENTSEvaluatingLarge2024}{2024}){]} investigated
LLM behavior in the popular game Among Us, revealing capabilities in
understanding complex game dynamics and successfully navigating roles
involving deception and cooperation. Similarly, Du et al.~{[}Du,
Rajivan, and Gonzalez
(\citeproc{ref-duLargeLanguageModels2024}{2024}){]} examined collective
problem-solving scenarios, finding that LLM agent groups showed
increased complexity in their interactions, more frequent disagreements,
and generally positive exchanges compared to human groups. Piatti et
al.~{[}Piatti et al.
(\citeproc{ref-piattiCooperateCollapseEmergence2024}{2024}){]}, through
the GovSim environment, studied how AI societies managed collective
resources, demonstrating that LLM agents effectively balanced ethical
considerations, strategic planning, and negotiation, further supporting
the idea of their advanced cooperative and strategic capabilities.

Within the specific context of Werewolf, several studies have addressed
the use of LLMs to enhance gameplay. Xu et al.~{[}Xu et al.
(\citeproc{ref-xuLanguageAgentsReinforcement2024}{2024}){]} developed
LLM agents that leverage deductive reasoning and reinforcement learning
to optimize decision-making and gameplay strategy, outperforming
existing methods. Meanwhile, Bailis et al.~{[}Bailis, Friedhoff, and
Chen (\citeproc{ref-bailisWerewolfArenaCase2024}{2024}){]} introduced
the Werewolf Arena, a framework employed in our current study. However,
despite these advancements, previous research has not explicitly
examined or compared LLM-driven Werewolf gameplay to authentic human
interactions and strategies. Our paper specifically addresses this gap
by directly comparing human gameplay, sourced from the Werewolf Among Us
dataset, with simulated gameplay generated by LLMs, providing novel
insights into differences and similarities in persuasion strategies and
social reasoning between humans and AI.

@@

Agents participated in gameplay through a graphical user interface
(GUI), depicted in Figure~\ref{fig-wwa-gui}. The GUI displays the game
state, including player roles, actions, and current discussion rounds,
enabling monitoring of the gameplay progression.
